---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
  html_notebook: default
---


```{r}
library(tidyverse)
all_counts <- readRDS(file= "Data/pc_count_matrix.rds")
avg_counts <- readRDS(file="Data/avg_pc_count_matrix.rds")
diff_counts <- readRDS(file="Data/diff_pc_count_matrix.rds")
meta <- suppressMessages(readr::read_tsv("Data/complete_meta_data.tsv"))
meta <- meta %>% mutate(tissue_type=replace(tissue_type, tissue_type=="embryonic facial prominence", "facial prominence"))
meta <- meta %>% mutate(tissue_type=replace(tissue_type, tissue_type=="skeletal muscle tissue", "muscle tissue"))
meta <- meta %>% mutate(tissue_type=replace(tissue_type, tissue_type=="skeletal muscle tissue", "muscle tissue"))
tissue_types <- unique(meta %>% pull(tissue_type))
collapsed_dev_counts <- matrix(data=NA, nrow=nrow(all_counts),ncol=length(tissue_types))
walker <- 1
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  tissue_subset_count <- all_counts[,tissue_subset_ids]
  tissue_avg <- rowMeans(tissue_subset_count)
  collapsed_dev_counts[,walker] <- tissue_avg
  walker <- walker + 1
}
colnames(collapsed_dev_counts) <- tissue_types
rownames(collapsed_dev_counts) <- rownames(all_counts)
TFs<- c("Ascl1", "Hes1", "Neurod1", "Mecp2", "Mef2c", "Runx1", "Tcf4", "Pax6")
```


Investigate the most highly expressed gene (sum of all stages) for each tissue type
```{r}
maxes_list <- list()
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_sums <- apply(all_counts[,tissue_subset_ids],1, sum)
  maxes_list[tissue] <- names(all_sums)[which.max(all_sums)]
}
for (name in names(maxes_list)){
  print(paste0("The most highly expressed protein in ", name, " : ", maxes_list[name]))
}
```
Same thing as above but excluding the mitochondria
```{r}
maxes_list <- list()
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_sums <- apply(all_counts[,tissue_subset_ids],1, sum)
  all_sums <- all_sums[!sapply(names(all_sums), startsWith, "mt-")]
  maxes_list[tissue] <- names(all_sums)[which.max(all_sums)]
}
for (name in names(maxes_list)){
  print(paste0("The most highly expressed protein in ", name, " : ", maxes_list[name]))
}
```
Top 5 Universally highly expressed protein across all tissues regardless of stages (without processing)
```{r}
##Collapse dev_stages
collapsed_matrices <- matrix(data=NA, nrow = nrow(all_counts), ncol = length(tissue_types))
rownames(collapsed_matrices) <- rownames(all_counts)
walker <- 1
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_avgs <- apply(all_counts[,tissue_subset_ids],1, mean)
  collapsed_matrices[,walker] <- all_avgs
  walker <- walker + 1
}
all_sums <- apply(collapsed_matrices,1, sum)
for (i in 1:10){
  print(paste0("The highests expressed protein acrossed all tissue in order ",i ,": ", names(sort(all_sums,decreasing=TRUE))[i]))
}


```
Expression profile of the highest expressed protein across all tissues Hba-a2
Extract only Hba-a2 information
```{r}
library(ggplot2)
target_protein <- "Hba-a2"
unique_dev_stages <- unique(meta %>% pull(dev_stage))
Hba_matrix <- matrix(data=NA, nrow=length(tissue_types), ncol=3)
colnames(Hba_matrix) <- c("count", "dev_stage", "tissue_type")
Hba_frame <- data.frame(Hba_matrix)
walker <- 1
for (exp_id in names(avg_counts[target_protein,])){
    Hba_frame[walker, "count"] <- avg_counts[target_protein, exp_id]
    Hba_frame[walker, "dev_stage"] <- meta %>% filter(id==exp_id) %>% pull(dev_stage)
    Hba_frame[walker, "tissue_type"] <- meta %>% filter(id==exp_id) %>% pull(tissue_type)
    walker <- walker + 1
}
Hba_frame$count <- log2(Hba_frame$count) 
```
Plot the extracted information 
```{r fig.height=8, fig.width=16}
library(ggrepel)
make_label <- function(row){
  if (row['dev_stage'] == max(Hba_frame %>% filter(tissue_type==row['tissue_type']) %>% pull(dev_stage))){
    #print(max(Hba_frame %>% filter(tissue_type==tissue) %>% pull(dev_stage)))\
     return(as.character(row['tissue_type']))
  }else {

    return(NA_character_)
  }
}
Hba_frame[,"label"] <- Hba_frame %>% apply(1,make_label)
# Hba_frame <- Hba_frame %>%
#   mutate(label = if_else(dev_stage == max(Hba_frame %>% filter()), as.character(tissue_type), NA_character_))
ggplot(Hba_frame, aes(x=dev_stage, y=count, group=tissue_type, colour=tissue_type)) + geom_line() + geom_point() + expand_limits(x= c(4, 10)) +
  # geom_text_repel(
  #   aes(label = gsub("^.*$", " ", label)),
  #   # This will force the correct position of the link's right end.	

  #   segment.curvature = -0.1,
  #   segment.square = TRUE,
  #   segment.color = 'grey',
  #   box.padding = 0.1,
  #   point.padding = 0.6,
  #   nudge_x = 1,
  #   nudge_y = 1,
  #   force = 15,
  #   hjust = 0,
  #   direction = "y",
  #   na.rm = TRUE,
  #   xlim = c(5, 10),
  #   ylim = c(0, 150000),
  # ) +
  geom_text_repel(data = . %>% filter(!is.na(label)),
                  aes(label = paste0("  ", label)),
                  #segment.alpha = 0, ## This will 'hide' the link
                  segment.curvature = -0.1,
                  segment.square = TRUE,
                  # segment.color = 'grey',
                  box.padding = 0.1,
                  point.padding = 0.6,
                  nudge_x = 1,
                  nudge_y = 1,
                 force = 0.5,
                  hjust = 0,
                  direction="y",
                  na.rm = TRUE, 
                  xlim = c(5, 10),
                  ylim = c(10,17),
)                   +
  ylab("counts(log2)") + ggtitle("Expression pattern of Hba-a2 across tissues") + theme(plot.title = element_text(hjust = 0.5))
```
Top 5 Universally highly expressed protein across all tissues regardless of stages (by percentage in tissues)
```{r}
##Collapse dev_stages
collapsed_matrices <- matrix(data=NA, nrow = nrow(all_counts), ncol = length(tissue_types))
rownames(collapsed_matrices) <- rownames(all_counts)
walker <- 1
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_avgs <- apply(all_counts[,tissue_subset_ids],1, mean)
  percent_counts <- all_avgs/sum(all_avgs)
  collapsed_matrices[,walker] <- percent_counts
  walker <- walker + 1
}
all_sums <- apply(collapsed_matrices,1, sum)
for (i in 1:10){
  print(paste0("The highests expressed protein acrossed all tissue in order ",i ,": ", names(sort(all_sums,decreasing=TRUE))[i]))
}

```
Top 5 Universally highly expressed protein across all tissues regardless of stages (by ranking)
```{r}
##Collapse dev_stages
collapsed_matrices <- matrix(data=NA, nrow = nrow(all_counts), ncol = length(tissue_types))
rownames(collapsed_matrices) <- rownames(all_counts)
walker <- 1
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_avgs <- apply(all_counts[,tissue_subset_ids],1, mean)
  all_ranking <- rank(all_avgs)
  collapsed_matrices[,walker] <- all_ranking
  walker <- walker + 1
}
all_sums <- apply(collapsed_matrices,1, sum)
for (i in 1:10){
  print(paste0("The highests expressed protein acrossed all tissue in order ",i ,": ", names(sort(all_sums, decreasing = TRUE))[i]))
}

```

expression distribution acrossed different tissues (box plot)
```{r, fig.height=8, fig.width=16}
library(cowplot)
all_counts_melted <- data.frame(all_counts) %>% gather("id", "count")
all_counts_melted <- all_counts_melted %>% inner_join((meta %>% select(id, tissue_type, dev_stage)), by="id")
  ggplot(data=all_counts_melted, aes(x=dev_stage, y=log2(count+1))) + geom_boxplot() +  facet_wrap(~ tissue_type, ncol=3) + theme_cowplot(12) + geom_hline(yintercept=median(log2(all_counts+1)), colour="red")

```
expression distribution acrossed different tissues (histogram)
```{r fig.height=8, fig.width=16, warning=FALSE}
all_counts_melted <- data.frame(all_counts) %>% gather("id", "count")
all_counts_melted <- all_counts_melted %>% inner_join((meta %>% select(id, tissue_type, dev_stage)), by="id")
  suppressWarnings(ggplot(data=all_counts_melted, aes(x=log2(count+1), fill=dev_stage)) + geom_histogram(binwidth = 0.5) +  facet_wrap(~ tissue_type, ncol=3) + theme_cowplot(12))

```

expression distribution acrossed different tissues (histogram log transformed)
```{r fig.height=8, fig.width=16, warning=FALSE}
all_counts_melted <- data.frame(all_counts) %>% gather("id", "count")
all_counts_melted <- all_counts_melted %>% inner_join((meta %>% select(id, tissue_type, dev_stage)), by="id")
  suppressWarnings(ggplot(data=all_counts_melted, aes(x=log2(count+1), y=..density.. ,fill=dev_stage)) + geom_histogram(binwidth = 0.5) +  facet_wrap(~ tissue_type, ncol=3) + theme_cowplot(12))
```
expression distribution acrossed different tissues (box plot collapsed dev_stages)
```{r, fig.height=8, fig.width=16}
melted_collapsed_dev <- pivot_longer(data.frame(collapsed_dev_counts), cols=(everything()), names_to = "tissue_type", values_to = "count")
ggplot(data=melted_collapsed_dev, aes(y=log2(count+1))) + geom_boxplot() +  facet_grid(. ~ tissue_type) + theme_cowplot(12) + theme(
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) + xlab("tissue_types") + geom_hline(yintercept=median(log2(all_counts+1)), colour="red")

```
expression distribution acrossed different tissues (histogram collapsed by dev_stage)
```{r fig.height=8, fig.width=16, warning=FALSE}
  suppressWarnings(ggplot(data=melted_collapsed_dev, aes(x=log2(count+1), fill=tissue_type, y=..density..)) + geom_histogram(binwidth = 0.5) + theme_cowplot(12))  
```
expression distribution acrossed different tissues (freq_plot collapsed by dev_stage)
```{r fig.height=8, fig.width=16, warning=FALSE}

  ggplot(data=melted_collapsed_dev, aes(x=log2(count+1), y=..density.., colour=tissue_type)) + geom_freqpoly(binwidth=0.5) + theme_cowplot(12) + scale_x_continuous(limits=c(0, 20))

```
Mean-variance
```{r fig.height=8, fig.width=16}
library(ggrepel)
all_counts_mean_var <- all_counts
all_counts_mean_var <- cbind(all_counts_mean_var, mean=apply(all_counts,1,mean))
all_counts_mean_var <- cbind(all_counts_mean_var, variance=apply(all_counts,1,var))
ggplot(data=data.frame(all_counts_mean_var), aes(x=log2(mean+1), y=log2(variance+1))) + geom_point() + theme_cowplot(12) + 
  geom_text_repel(data=data.frame(all_counts_mean_var) %>% rownames_to_column(var="rowname") %>% slice_max(variance, n=10), aes(x=log2(mean+1), y=log2(variance), label=rowname, colour="red")) +
  geom_text_repel(data=data.frame(all_counts_mean_var) %>% rownames_to_column(var="rowname") %>% slice_max(mean, n=10), aes(x=log2(mean+1), y=log2(variance), label=rowname, colour="blue")) + scale_colour_discrete(name = "Top elements", labels = c("Variance", "Mean"))
```
Mean and variance all together but hue by organism type
```{r fig.height=8, fig.width=16, warning=FALSE}

collapsed_matrices <- data.frame(mean=double(), variance=double(), tissue=character(), row=character())
colnames(collapsed_matrices) <- c("mean", "variance", "tissue")
for (tissue in tissue_types){
  tissue_subset_ids <- meta %>% filter(tissue_type == tissue) %>% pull(id)
  all_means <- apply(all_counts[,tissue_subset_ids],1, mean)
  all_variance <- apply(all_counts[,tissue_subset_ids],1, var)
  temp_matrix <- data.frame(mean=all_means, variance=all_variance, tissue) %>% rownames_to_column("rowname")
  collapsed_matrices <- rbind(collapsed_matrices, temp_matrix)
}
ggplot(data=collapsed_matrices, aes(x=log2(mean+1), y=log2(variance+1), colour=tissue)) + geom_point()+ theme_cowplot(12) + 
  geom_text_repel(data=collapsed_matrices %>% slice_max(variance+mean, n=15), aes(x=log2(mean+1), y=log2(variance+1), label=rowname, colour=tissue)) 
  #geom_text_repel(data=collapsed_matrices %>% rownames_to_column("row") %>% slice_max(mean, n=10), aes(x=log2(mean+1), y=log2(variance+1), label=row, colour=tissue))
  
  #scale_colour_discrete(name = "Top elements", labels = c("Variance", "Mean"))
```
Variability across tissue types 
```{r fig.height=8, fig.width=16, warning=FALSE}



ggplot(data=collapsed_matrices, aes(x=log2(mean+1), y=log2(variance+1))) + geom_point() +  facet_wrap(~ tissue, ncol=3) + theme_cowplot(12) 
```
Zooming in on highly epxressed gene (>1025)
```{r fig.height=8, fig.width=16}
ggplot(data=collapsed_matrices  %>% filter(log2(mean+1) >=10), aes(x=log2(mean+1), y=log2(variance+1))) + geom_point() +  facet_wrap(~ tissue, ncol=3) + theme_cowplot(12) 
```
Distribution of the TFs
```{r fig.height=8, fig.width=16}
TFCounts <- data.frame(all_counts[TFs,])
TFCounts <- TFCounts %>% rownames_to_column("rowname")
TFCounts <- TFCounts %>% gather("id", "counts", -rowname)
TFCounts <- TFCounts %>% left_join((meta %>% select(tissue_type, dev_stage, id)), by="id")

ggplot(data=TFCounts, aes(x=counts)) + geom_histogram(bins = 100) + facet_grid(~  rowname) + theme_cowplot(12)
```
Distribution of the TFs (Log transformed)
```{r fig.height=8, fig.width=16}
TFCounts <- data.frame(all_counts[TFs,])
TFCounts <- TFCounts %>% rownames_to_column("rowname")
TFCounts <- TFCounts %>% gather("id", "counts", -rowname)
TFCounts <- TFCounts %>% left_join((meta %>% select(tissue_type, dev_stage, id)), by="id")

ggplot(data=TFCounts, aes(x=log2(counts+1))) + geom_histogram(bins = 100) + facet_grid(~  rowname) + theme_cowplot(12)
```
Distribution of the TFs (Log transformed)
```{r fig.height=8, fig.width=16}
TFCounts <- data.frame(all_counts[TFs,])
TFCounts <- TFCounts %>% rownames_to_column("rowname")
TFCounts <- TFCounts %>% gather("id", "counts", -rowname)
TFCounts <- TFCounts %>% left_join((meta %>% select(tissue_type, dev_stage, id)), by="id")

ggplot(data=collapsed_matrices %>% filter(rowname %in% TFs), aes(x=log2(mean+1), y=log2(variance+1), colour=tissue)) + geom_point()+ theme_cowplot(12) + 
  geom_text_repel(data=collapsed_matrices %>% filter(rowname %in% TFs) %>% slice_max(variance+mean, n=15), aes(x=log2(mean+1), y=log2(variance+1), label=rowname, colour=tissue)) 

```
Box plots for TFs across tissues
```{r fig.height=8, fig.width=16}
Tcf4Counts <- TFCounts %>% filter(rowname=="Tcf4")

ggplot(data=TFCounts, aes(y=counts, x=tissue_type)) + geom_boxplot() + theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) + facet_wrap(~ rowname, scales = "free") + geom_hline(yintercept=mean(TFCounts$counts), colour="red")
```
Tcf4 seems to be involved in neuronal differentiation so it makes sense for it to be highly expressed in the brains/neural tube but no so much in liver
Some interesting observation, Mecp2 is basically not epxressed at all same as Runx1(except in thymus), and Tcf4 is one of the most active

Expression across dev_stage and tissues
```{r fig.height=8, fig.width=16}

ggplot(data=TFCounts, aes(x=dev_stage, y=counts, colour=rowname)) + geom_point() + theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) + facet_wrap(~ tissue_type, scales = "free") + scale_colour_discrete(name = "Gene")
```
Expression across dev_stage and tissues (log transformed)
```{r fig.height=8, fig.width=16}

ggplot(data=TFCounts, aes(x=dev_stage, y=log2(counts+1), colour=rowname)) + geom_point() + theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) + facet_wrap(~ tissue_type) + scale_colour_discrete(name = "Gene")
```
histograms of expresssion distribution by tissues
```{r fig.height=8, fig.width=16}

ggplot(data=TFCounts, aes(x=log2(counts+1), fill=rowname)) + geom_histogram(bins=100) + theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1)) + facet_wrap(~ tissue_type, ncol=3,scales = "free") + scale_colour_discrete(name = "Gene")

```
cluster correlation heatmap
```{r fig.height=8, fig.width=16}
library(gplots)
library(RColorBrewer)
# Get some nicer colours
mypalette <- brewer.pal(11, "RdYlBu")
# http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3
morecols <- colorRampPalette(mypalette)
# Set up colour vector for celltype variable
col.cell <- (c("purple","orange")[meta$tissue_type])[1:(length(meta$tissue_type)/2)]
# Plot the 
avg_counts_tissue_name <- avg_counts
colnames(avg_counts_tissue_name) <- sapply(colnames(avg_counts), function(x){temp_row <- meta %>% filter(id==x) 
                                                                             return(paste0(temp_row %>% pull(tissue_type), " ",temp_row %>% pull(dev_stage)))})
heatmap.2(log2(avg_counts_tissue_name+1)[(order(apply(avg_counts,1,var), decreasing = TRUE))[1:50],], 
          col=rev(morecols(50)),
          trace="column", 
          main="Top 50 most variable genes across samples",
          ColSideColors=col.cell,
          scale = "row")
```

cluster correlation heatmap (only TFs)
```{r fig.height=8, fig.width=20}
library(gplots)
library(RColorBrewer)
# Get some nicer colours
mypalette <- brewer.pal(11, "RdYlBu")
# http://colorbrewer2.org/#type=sequential&scheme=BuGn&n=3
morecols <- colorRampPalette(mypalette)
# Set up colour vector for celltype variable
col.cell <- (c("purple","orange")[meta$tissue_type])[1:(length(meta$tissue_type)/2)]
# Plot the heatmap
avg_counts_tissue_name <- avg_counts
colnames(avg_counts_tissue_name) <- sapply(colnames(avg_counts), function(x){temp_row <- meta %>% filter(id==x) 
                                                                             return(paste0(temp_row %>% pull(tissue_type), " ",temp_row %>% pull(dev_stage)))})
heatmap.2(log2(avg_counts_tissue_name+1)[TFs,], 
          col=rev(morecols(length(TFs))),
          trace="column", 
          main="TF count heat map across tissues",
          ColSideColors=col.cell,
          scale="row")
```
Midbrain cluster correlation heatmap (only TFs)
```{r fig.height=8, fig.width=20}
# Set up colour vector for celltype variable
col.cell <- (c("purple","orange")[meta$tissue_type])[1:(length(meta %>% filter(tissue_type=="midbrain") %>% pull(id)))]
# Plot the heatmap
midbrain_id <- meta %>% filter(tissue_type == "midbrain") %>% pull(id)
midbrain_counts <- all_counts[,midbrain_id]
colnames(midbrain_counts) <- sapply(colnames(midbrain_counts), function(x){temp_row <- meta %>% filter(id==x) 
                                                                             return(paste0(temp_row %>% pull(dev_stage)))})
heatmap.2(log2(midbrain_counts+1)[TFs,], 
          col=rev(morecols(length(TFs))),
          trace="column", 
          main="TF count heat map across Midbrain",
          ColSideColors=col.cell,
          scale="row")
```
Hindbrain cluster correlation heatmap (only TFs)
```{r fig.height=8, fig.width=20}
# Set up colour vector for celltype variable
col.cell <- (c("purple","orange")[meta$tissue_type])[1:(length(meta %>% filter(tissue_type=="hindbrain") %>% pull(id)))]
# Plot the heatmap
hindbrain_id <- meta %>% filter(tissue_type == "hindbrain") %>% pull(id)
hindbrain_counts <- all_counts[,hindbrain_id]
colnames(hindbrain_counts) <- sapply(colnames(hindbrain_counts), function(x){temp_row <- meta %>% filter(id==x) 
                                                                             return(paste0(temp_row %>% pull(dev_stage)))})
heatmap.2(log2(hindbrain_counts+1)[TFs,], 
          col=rev(morecols(length(TFs))),
          trace="column", 
          main="TF count heat map across Hindbrain",
          ColSideColors=col.cell,
          scale="row")
```






Cluster by correlation
PCA by cell type
Check if using TPM







variability (across tissues, within tissues as well) mean-variance is the ranking similar across tissues
```{r fig.height=8, fig.width=16}
##Remove zero variance 
all_counts_drop_var <- (t(all_counts))[, which(apply(t(all_counts),2,var) !=0)]
all_counts.pca <- prcomp(all_counts_drop_var, center = TRUE,scale. = TRUE)
library(ggfortify)
autoplot(all_counts.pca, data=meta, colour="dev_stage") + theme_cowplot(12)
```













